\cursedforest\ - A Random Forest Implementation for ``Big'' and ``Wide'' Data \\ Supplementary Information

### 2017-01-17 doc/publications/plos16/plos16.tex
add to BMD section

%% Additional packages
\usepackage{footmisc}
\usepackage{multirow}
%\usepackage[T1]{fontenc}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
We consider the problem of feature discovery on a genomic dataset for a specific phenotype.
In this case we investigate the ability to retrieve variants in the same loci as genes previously demonstrated to
be associated with a high or low bone mass density (BMD) and rank these variants as being "important".
We build a random forest model on this data, and produce a list of the top 5,000 contributing features, ranked by their
Gini impurity. We apply the Mann–Whitney U test to these fetures and observe a p-value of x, supporting our hypothesis
that features in close proximity to the known BMD genes are contributing more information to our classification model.

However, we also observe that from the set of highest-ranked features---for example, the top ten---not all of them exist
within close proximity to known BMD genes.

One example is COLEC10. Featuring a transcription factor binding site for FOXL1, COLEC10 is also reported as containing a SNP
associated with BMD \cite{Liu2008}.

We also frequently see highly-ranked variants in chromosome 22q11

@@

\begin{figure}[tbhp] 
 \begin{adjustwidth}{-1.00in}{0in}
    \centering
    \includegraphics[totalheight=6cm]{./figs/phase.png} 
    \caption{{\bf The phase chage diagram.}}
    \label{figure:phase-diagram-equivalence.png} 
    \vspace{4ex}
  \end{adjustwidth}
\end{figure}

@@

\bibliography{./CursedForest.bib}

>>>>

%% Additional packages
\usepackage{footmisc}
\usepackage{multirow}
\usepackage[T1]{fontenc}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Section 2: Feature Selection on real data 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{\cursedforest\ selects previously demonstrated bone-density variants}
We investigate the ability of using a random forest model we build on genomic data to identify variants related to a
specific phenotype.
In this case we aim to identify variants associated with Bone Minderal Density (BMD). We build a model on the BMD dataset
(see datasets) using a binary high or low BMD value as the label.
From this model, we output a list of the top 10,000 contributing variants, ranked by their Gini impurity. The variant with
the highest Gini impurity---and hence, importance---is ranked 1, and the variant with the lowest Gini impurity is ranked 10,000.
Then, using the 26 known and 6 novel BMD genes from Duncan et al., we compile a list of variants that are present in or near one of these
32 genes.

We then apply the Mann–Whitney U test to this list to verify that from our list, variants from known BMD loci are ranked
higher than variants from elsewhere in the genome.
We observe a p-value of XXXX, supporting our hypothesis that variants in close proximity to the known BMD genes are
contributing more information to our classification model.

However, we also observe that from the set of highest-ranked features---for example, the top ten---not all of them exist
within close proximity to known BMD genes. We investigate the literature to determine the plausibility of these variants
contributing to BMD.

One gene is COLEC10 (collectin 10). Featuring a transcription factor binding site for FOXL1, COLEC10 is also reported as
containing variants associated with BMD \cite{Kemp2014,Liu2008}. However there currently exists no information on how this gene
may contribute to BMD.

We also frequently see highly-ranked variants in chromosome 22q11. Stagi2010 observe that individuals with a microdeletion
in this region have a significant in bone mass\cite{Stagi2010}. In this same region, we also observe variants present in
PRODH (proline dehydrogenase). PRODH catalyzes the first step in proline degredation, and as proposed by Phang et al., proline
is made available by collagen degradation \cite{Phang2008}.

DCDC5 (Doublecortin Domain Containing 5) is one of the 26 known genes, however it is one of the genes that Duncan et al.
failed to replicate in their study. DCDC5 does appear in our list of genes, however the first occurance of a variant in
or near this gene has a rank of 9,667 out of 10,000. Also, this gene is not highly expressed in bone, and any mechanisms
by which this gene might regulate BMD are unclear \cite{Thakker2012}. Although variants near DCDC5 are associated with
BMD \cite{Rivadeneira2009}, it is questionable as to whether the gene itself contributes to BMD.

@@

\begin{figure}[tbhp] 
 \begin{adjustwidth}{-1.00in}{0in}
    \caption{\textbf{The Donoho-Tanner phase chage diagram.}}
    \centering
    \includegraphics[totalheight=6cm]{./figs/phase.png} 
    \label{figure:phase-diagram-equivalence.png} 
    \vspace{4ex}
  \end{adjustwidth}
\end{figure}

@@

\bibliography{CursedForest}
